---
title: "Extended Abstract: Extended Early Adulthood and the Unchanging Probability of Opinion Change"
author:
  - Kevin Kiley, North Carolina State University
  - Nicolas Restrepo, University of California, Davis
fontsize: 12pt
output:
  pdf_document: default
  html_document: default
header-includes:
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \def\linenumberfont{\normalfont\tiny\sffamily}
bibliography: efw_bib.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(tidyverse)
library(brms)
library(patchwork)
theme_set(jtools::theme_apa())
```

## Motivation

Adolescence and young adulthood are typically identified in the social science literature as "impressionable years" -- a stage of the life course in which people are more susceptible to broad social currents; when their attitudes are more malleable; and when their core dispositions, identities, and opinions are formed [@alwin1991]. A broad body of research finds that rates of opinion change tend to peak during the ages of 14 to 25, with opinions becoming more stable once people move into middle age [@sears1999; @bartels2014; @ghitza2022]. And this recurring empirical pattern is central to broader dynamics around cohortization that are frequently invoked in larger models of social change [@ryder1965; @mannheim1952; @vaisey2016].

However, why people in these life stages appear to show more frequent opinion change relative to older people is unclear. Work in biology, psychology, and cultural evolution suggests that peoples' cognition changes in ways that might make them more resistant to new information as they age. Explanations in sociology, in contrast, tend to focus on the unique social features of these life stages -- including high rates of mobility between social contexts and participation in educational institutions -- that shape how people are exposed to new sources of information and their ability to update opinions in the face of this information [@visser2004]. While both mechanisms likely play a role in explaining the higher rates of opinion change in this life stage, the relative contributions of each are unclear.

In this paper, we leverage changes in the timing, ubiquity, and standardization of life course transitions over the second half of the 20th Century [@buchmann1989; @brückner2005] to adjudicate the relative contributions of developmental and social mechanisms to the heightened rates of opinion change we observe in young people. Since the 1950s, the median age of marriage and childbirth have each been pushed back by about 10 years for both men and women [@eickmeyer2017; @manning2014], rates of college-going have increased dramatically, and people make more transitions between workplaces and between the workforce and education [@brückner2005]. These shifts have led to a dramatically expanded window of participation in educational institutions, geographic and social mobility, and freedom from various social obligations and expectations, all of which are assumed to facilitate opinion change [@rosenfeld2005]. If the heightened rates of opinion change we observe in young adults are principally due to the social structure of this life stage, then we should observe changes in the rates of opinion change in this group over time.

## Data & Analytic Strategy

Drawing on four nationally representative panel surveys, dating from the mid-1950s through the early 2000s, we compare multiple different measures of opinion change in general political opinions and partisan identification of early adults (people under 30) and the rest of the population. We draw on four different panel surveys of the American adult population: the 1956-60, 1972-76, and 1992-96 American National Election Study panels, and the 2006-10 General Social Survey panel. All four surveys are nationally representative samples of non-institutionalized adults in the United States. All panels also interviewed survey respondents three times over a four-year window, making the duration of each survey window comparable. All four panels also include comparable questions about general sentiments toward government action, a measure of partisan identification, and measures of life course transitions such as marriage and childbearing.

We focus on questions about general political dispositions that were asked in all three waves of each panel, with a total of 29 questions, with between 6 and 8 questions in each panel. These questions tend to focus on the national government's role in the economy, the government's obligation to ensure equal opportunity or outcomes, and the government's role in helping minorities, as well as self-placement on a liberal to conservative scale. We also include a seven-point measure of partisan identification that is asked similarly in each panel. No other questions are directly comparable across all of the panels, though some questions are asked in comparable ways across at least two panels. Rather than adopt a single approach to quantify the rates of change across questions, we use a variety of measures of "change." We calculate the absolute difference between respondents' opinions at time 3 and their response at time 1 ("Total Distance"), the standard deviation of responses at all three waves ("S.D."), and construct indicators for whether they cross or select the scale midpoint at any point over the three waves ("Cross"), whether they finish on a different side of the scale from where they started ("Cross+Stay"), and whether they change more than two positions on the five- or seven-point scale that they are measured on ("Change \>2"). Each of these measures has different strengths and weaknesses, though none perfectly captures an error-free measure of change.

For each definition of change, we regress that outcome on an indicator variable for whether people are aged 30 or less. For the continuous outcomes, we use ordinary least squares regression and for the dichotomous outcomes we use logistic regression. In later analyses, we include covariates for marital status, whether the respondent has children, and their highest level of education to evaluate whether these life-course transitions account for (or obscure) gaps between age groups over time. This approach produces an extremely large number of models, and will likely produce statistically significant coefficients at the $p < .05$ level for a substantial proportion of coefficients simply by chance. Our goal is not to interpret any single coefficient but to interpret overall patterns if they emerge. To that end, we principally focus on comparing plots of many models rather than evaluating specific questions.

## Results

Figure 1 plots the predicted opinion change (for scalar outcomes: total distance and standard deviation) or predicted probability of opinion change (all other measures) for each of these questions, for respondents aged 30 and older (red dots) and for respondents under 30 (blue dots). Each column represents a different measure of "change" as outlined above, and the vertical axis is separated into the different panels, with questions from the oldest panel (the 1956-60 ANES panel) at the top and questions from the most recent panel (the 2006-10 GSS) at the bottom. If early adults have become more likely to make opinion changes over time relative to people over 30, we should see points diverge as we move down the figure.

```{r, include=TRUE, fig.height=8, fig.align="center", fig.cap=c("Predicted change distance or probability of change"), echo=FALSE}
knitr::include_graphics("./predicted_change_values.pdf")
```

While there are idiosyncratic questions in each panel, broadly speaking, there are a number of consistent patterns across the panels. First, there are high rates of opinion instability in the population, with the proportion of respondents crossing or selecting the scale midpoint ("Cross") often exceeding 80 percent for both early adults and the rest of the population. At the same time, little of this change is durable, with the proportion of respondents ending the survey on a different side of the scale ("Cross+Stay") often being less than 10 percent of the population. In other words, respondents are highly inconsistent in reporting their general political beliefs and tend to report opinions near the scale mid-point. This pattern is consistent with several decades of work on the stability of Americans' political opinions [@converse1964; @zaller1992]

Second, the population as a whole does not appear to become either more or less consistent on these general political issues over time. While it is possible that variation in which questions are included in each panel obscure an overall trend, there is no obvious shift in stability or instability over time in the measures we looked at. An average of 57 percent of respondents over 30 switched sides of the scale to each question in the 1950s, and about 60 percent of respondents over 30 switched sides of the scale to each question in the 2000s. That is remarkable stability despite significant shifts in the partisan landscape, the issues at the center of American politics, and the composition of the population.

Third, there is no consistent age-based pattern of opinion stability, especially over time. On most questions, in most panels, the rates of opinion change between early adults (under 30) and the rest of the population are statistically indistinguishable. This is not simply an issue of statistical power. Across all issues, young respondents are about as likely to have lower rates of opinion change than older people as higher rates of opinion change, and these differences are frequently only one or two percentage points. When we create composite scales from all issue questions in a panel by rescaling and averaging all questions so that higher values indicate more conservative responses, an approach designed to minimize measurement error in issue positions [@ansolabehere2008], we find no statistically significant difference between people under 30 and people 30 and older in any of the panels (not shown). This is somewhat surprising, given the general expectation that young people would be more likely to change opinions.  

At the same time, there are some observable differences between young and old respondents on a small number of questions that do reinforce the general expectation. On partisan identification, the one question included in all panels, respondents under 30 are consistently more likely than older respondents to change sides of the scale. But we do not see a clear trend over time in this difference. This question also shows lower overall rates of inconsistency across both young people and old people, suggesting it behaves quite differently from the other kinds of opinions analyzed here.

None of this should imply that young people do not have substantively different opinions (they likely do, but that is beyond the scope of this analysis). But what we can say is that they are not any more or less stable by the metrics we outlined above. There are only a handful of issues where we observe a consistent pattern across most measures of attitude change: partisan identification in most panels, the issue of busing in the 1970s, and general views on government spending in the 1990 panel.

Further analyses (not included here) explore whether life-course transitions obscure or exacerbate age-based differences. We find little relationship between marital status and the likelihood of opinion change in any panel. While higher levels of education are associated with less change over time in the later panels, we find that including it in a model with age does not affect the age coefficient in any meaningful way.

## Discussion

Despite significant shifts in the timing of life course events and the broad social structure of early adulthood from the mid-1950s to today, there is little change in the rates at which people between the ages of 18 and 30 undergo major changes of political opinions. We suggest four possible explanations for this aggregate stability. First, and we believe most plausibly, the patterns observed in this analysis broadly suggest that there is something developmental that explains broad patterns in attitude instability early in the life course. Second, it is possible that shifts in social structure push in opposite directions to produce aggregate stability both in the rate at which respondents over 30 make changes of opinion and the difference between younger and older respondents. However, these shifts would have to be unrelated to life-course transitions like completing education and getting married, which are unrelated to the difference between young and old respondents. Third, it is possible that the features that explain stability are still social in nature, but are rooted in factors that have not undergone significant changes since the 1950s, though it is unclear what those might be. Finally, it might be the case that issue turnover across the panels explains the lack of a trend. It could be the case that panels select questions that produce high amounts of opinion instability in the general population and that obscure differences between young and old people. However, because we observe no clear trend in the one question that is asked in all panels, partisan identification, we think this is unlikely.

## References
